PCA

### 2019-11-11 analysis/11_pcalda.Rmd
now with char and manip

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
library(Rtsne)
library(viridis)

@@

# custom theme for plots
mytheme <- function(){
  theme_minimal(base_size = 12) + 
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "transparent", colour = NA),
          plot.background = element_rect(fill = "transparent", colour = NA),
          legend.title = element_blank())
}
```

@@

# combined pca tsne

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
library(Rtsne)
library(viridis)
library(forcats)

@@

source("../R/themes.R")  
```

@@

## all data

```{r charmanip}
# prep col data for all samples
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = alllevels2)
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
colData$tissue <- factor(colData$tissue, levels = tissuelevels)
row.names(colData) <- colData$V1

# prep count data for all samples
countData <- as.data.frame(pseudocounts)
countData$entrezid <- countData$X1
countData <- left_join(geneinfo, countData)
countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
    select(-row.names, -geneid, -entrezid, -X1) 
countData <- as.data.frame(countData)
row.names(countData) <- countData$Name
countData$Name <- NULL
countData$entrezid <- NULL
countData <- as.data.frame(t(countData))

# check ready for analysis
row.names(countData) == row.names(colData)

# pca
mypca <- prcomp(countData)
eig.val <- get_eigenvalue(mypca)
head(eig.val)

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
head(mypcadf)

all1 <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         #shape = colData$tissue,
                         #size = colData$sex,
                        shape = colData$study)) +
  geom_hline(yintercept=0, color = "grey") +
  geom_vline(xintercept=0, color = "grey") +
  geom_point(size = 2)  + theme_minimal(base_size = 6)  +
  mytheme() +
  theme(legend.title = element_blank(),
        axis.text = element_blank()) +
  #scale_alpha_manual(values = c(0.5,1)) +
  labs(x = "PC1: 44.6% of variance", y = "PC2: 29.8% of variance") +
  scale_color_manual(values = colorscharmaip) 
all1

all2 <- fviz_pca_var(mypca,  labelsize = 4 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
    geom_hline(yintercept=0, color = "grey") +
  geom_vline(xintercept=0, color = "grey") +
    theme( axis.text = element_blank())

# prep for tsne
euclidist <- dist(countData) # euclidean distances between the rows

tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=10, theta=0.5, dims=2)
tsne_df = as.data.frame(tsne_model$Y) 
plot(tsne_df$V1, tsne_df$V2)

ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
  labs(x = "tSNE 1", y = "tSNE 2")

ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
  labs(x = "tSNE 1", y = "tSNE 2")

ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = study, color = treatment)) +
  geom_point(size = 2) +
  mytheme() +
  labs(x = "tSNE 1", y = "tSNE 2") +
  scale_alpha_manual(values = c(0.5,1)) +
  scale_color_manual(values = colorscharmaip)
```

```{r transparentplots}
png('../figures/pca/all-1.png',width=600,height=600, units="px",bg = "transparent")
print(all1)
dev.off()

png('../figures/pca/all-2.png',width=600,height=600, units="px",bg = "transparent")
print(all2)
dev.off()
```
